---
title: "Take-home Exercise 2"
author: "FU YILIN"
---

# **Getting Started**

In this exercise, we will be exploring the Vast Challenge 2024 - Mini-Challenge 1

# **Introduction**

## Setting the scene

Sailor Shift is a global music icon who began her career on the island nation of Oceanus, rooted in the local genre of Oceanus Folk. From modest beginnings, she rose to international fame, initially performing with the all-female band Ivy Echoes (2023–2026), before achieving a solo breakthrough in 2028 with a viral single that topped global charts — an unprecedented milestone for Oceanus Folk.

Since then, Sailor has released highly successful albums almost annually, explored Indie Pop and Indie Folk, and collaborated widely. Despite branching out, she often returns to Oceanus Folk, playing a major role in expanding its influence globally. Her dual passion for spreading Oceanus Folk and uplifting lesser-known artists is reflected in her frequent collaborations.

Sailor’s fame has also brought attention to her former Ivy Echoes bandmates, together, they’ve helped elevate Oceanus Folk into a vibrant and respected genre, inspiring a new generation of Oceanus artists.

As Sailor returns home in 2040, journalist Silas Reed is preparing a retrospective piece, “Oceanus Folk: Then-and-Now”, and has compiled a large knowledge graph of artists, albums, songs, and relationships.

## Key Tasks

In this Exercise, we will use the visualizations to develop a profile of what it means to be a rising star in the music industry.

1.  Visualize the careers of three artists. Compare and contrast their rise in popularity and influence.

2.  Using this characterization, give three predictions of who the next Oceanus Folk stars with be over the next five years.

# **Data Preparation**

## Load Packages

In the code chunk below, `p_load()` of **pacman** package is used to load the R packages into R environemnt.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts, 
               concaveman, ggforce, jsonlite, dplyr, SmartEDA, knitr, RColorBrewer, ggthemr,htmltools,patchwork, purrr, glue, kableExtra, fmsb, scales)
```

## Import Data

In the code chunk below, fromJSON() of jsonlite package is used to import MC1_graph.json file into R and save the output object.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
kg <- fromJSON("data/MC1_graph.json")
```

## Inspect Structure

In the code chunk below `str()` is used to reveal the structure of *kg* object.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
str(kg, max.level = 1)
```

## **Extracting the edges and nodes tables**

Next, `as_tibble()` of **tibble** package package is used to extract the nodes and links tibble data frames from *kg* object into two separate tibble data frames called *nodes_tbl* and *edges_tbl* respectively.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
nodes_tbl <- as_tibble(kg$nodes)
edges_tbl <- as_tibble(kg$links) 
```

# **Initial EDA**

In this code chunk below, ggplot2 functions are used the reveal the frequency distribution of *Edge Type* field of *edges_tbl*.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggthemr("dust") 

ggplot(data = edges_tbl,
       aes(y = `Edge Type`)) +
  geom_bar()
```

The code chunk below uses ggplot2 functions to reveal the frequency distribution of *Node Type* field of *nodes_tbl*.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ggthemr("dust") 

ggplot(data = nodes_tbl,
       aes(y = `Node Type`)) +
  geom_bar()
```

# **Creating Knowledge Graph**

## **Mapping from node id to row index**

Before we can go ahead to build the tidygraph object, it is important for us to ensures each id from the node list is mapped to the correct row number. This requirement can be achive by using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
id_map <- tibble(id = nodes_tbl$id,
                 index = seq_len(
                   nrow(nodes_tbl)))
```

## **Map source and target IDs to row indices**

Next, we will map the source and the target IDs to row indices by using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
edges_tbl <- edges_tbl %>%
  left_join(id_map, by = c("source" = "id")) %>%
  rename(from = index) %>%
  left_join(id_map, by = c("target" = "id")) %>%
  rename(to = index)
```

## **Filter out any unmatched (invalid) edges**

Lastly, the code chunk below will be used to exclude the unmatch edges.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
edges_tbl <- edges_tbl %>%
  filter(!is.na(from), !is.na(to))
```

## **Creating tidygraph**

Lastly, `tbl_graph()` is used to create tidygraph’s graph object by using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
graph <- tbl_graph(nodes = nodes_tbl, 
                   edges = edges_tbl, 
                   directed = kg$directed)
```

Confirm the output object is indeed in tidygraph format by using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
class(graph)
```

## **Visualising the knowledge graph**

In this section, we will use ggraph’s functions to visualise and analyse the graph object.

Several of the ggraph layouts involve randomisation. In order to ensure reproducibility, it is necessary to set the seed value before plotting by using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
set.seed(1234)
```

## **Visualising the whole graph**

In the code chunk below, ggraph functions are used to visualise the whole graph.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| 
ggraph(graph, layout = "fr") +
  geom_edge_link(alpha = 0.3, 
                 colour = "gray") +
  geom_node_point(aes(color = `Node Type`), 
                  size = 4) +
  geom_node_text(aes(label = name), 
                 repel = TRUE, 
                 size = 2.5) +
  theme_void()
```

Notice that the whole graph is very messy and we can hardy discover any useful patterns. In order to gain meaningful visual discovery, it is always useful for us to looking into the details, for example by plotting sub-graphs.

The following analysis will be proving in the article below.

::: title-box
# Oceanus Folk: Then-and-Now

*By Silas Reed \| Jun 1, 2040*
:::

# **🎶 Chapter 1**

## Charting the Ascent

### What Defines a Rising Star?

In the ever-shifting tides of the music industry, the phrase “rising star” is often tossed around—but what does it truly mean to earn that title?

To uncover the anatomy of a rising star, we turned to data: chart-topping hits marked as notable works served as the entry point. From there, we traced the artists responsible for these impactful songs and albums through *PerformerOf* connections in the creative network.

But rising stars aren’t defined by a single moment. For each artist in this cohort, we dug deeper—mapping every work they performed, whether it climbed the charts or not. The goal: to reveal a pattern of sustained output and influence. From these works, we extracted core attributes like *release year*, *genre*, and *notoriety year* (when a piece reached public acclaim), helping us track the arc of each artist’s rise.

Through this lens, a clear profile emerged. A true rising star is someone who:

-   Releases steadily, maintaining visibility with one or two new works every year.

-   Builds recognition across time, landing notable works not just once, but over multiple years.

-   Bridges experimentation and impact, often exploring diverse genres while resonating widely enough to reach the charts.

```{r}
#| code-fold: true
#| code-summary: "Show the code"

# Extract PerformerOf edge
graph_performers <- graph %>%
  activate(edges) %>%
  filter(`Edge Type` == "PerformerOf")

# Extract Performer nodes & works information
artist_influence <- graph_performers %>%
  activate(edges) %>%
  as_tibble() %>%
  left_join(nodes_tbl %>% select(id, name, `Node Type`, genre, notable, release_date, notoriety_date),
            by = c("target" = "id")) %>%
  rename(work_id = target,
         artist_id = source,
         work_name = name,
         work_type = `Node Type`) %>%
  left_join(nodes_tbl %>% select(id, name, stage_name), 
            by = c("artist_id" = "id")) %>%
  rename(artist_name = name) %>%
  filter(work_type %in% c("Song", "Album"),
         notable == TRUE) %>%
  select(artist_id, artist_name, stage_name, work_id, work_name, work_type, genre, 
         release_date, notoriety_date, notable)

# Export as CSV
write_csv(artist_influence, "artist_influence.csv")

#Convert "release_date" & "notoriety_date" as integer

influence <- artist_influence %>%
  mutate(
    release_year = as.integer(release_date),
    notoriety_year = as.integer(notoriety_date),
    is_notable = if_else(is.na(notoriety_date), FALSE, TRUE)
  )
```

### Constructing the Profile Criteria for Rising Stars

#### **Characteristic Traits (Rising Star Profile)**

-   Consistent Output: Releases at least 1–2 works per year to maintain active presence in the industry.

-   Influence Growth Across Years: Demonstrates sustained influence by having notable (charting) works across multiple years, not just a one-time breakout.

```{r}
#| label: rising-star-summary
#| code-fold: true
#| code-summary: "Show the code"
#| warning: false
#| message: false
# Step 1: Summarise artist performance metrics
artist_summary <- influence %>%
  group_by(artist_name) %>%
  summarise(
    total_works = n(),
    years_active = n_distinct(release_year),
    first_notoriety = if (all(is.na(notoriety_year))) NA_integer_ else min(notoriety_year, na.rm = TRUE),
    notable_works = sum(is_notable),
    notable_years = n_distinct(notoriety_year[!is.na(notoriety_year)])
  ) %>%
  ungroup() %>%
  mutate(
    avg_works_per_year = round(total_works / years_active, 1),
    rising_star_score = notable_works * 2 + total_works + notable_years * 1.5
  ) %>%
  arrange(desc(rising_star_score))

# Step 2: Display the top 10 rising stars in a formatted table
artist_summary %>%
  slice_head(n = 10) %>%
  kbl(
    caption = "Table 1. Top 10 Rising Stars",
    col.names = c("Artist Name", "Total Works", "Years Active", "First Chart Year",
                  "Notable Works", "Notable Years", "Avg. Works/Year", "Rising Star Score"),
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = "#f2f2f2")
```

This ranking highlights the most promising music artists based on their volume of releases, consistency of activity, and multi-year chart presence. **Kimberly Snyder**, **Yang Zhao**, and **Ping Tian** top the list, showing strong momentum and diverse musical influence.

# **🎶 Chapter 2**

## **Charting the Soundscape**

### How Rising Stars Shape Their Popularity Across Genres

In the ever-evolving landscape of music, genre isn’t just a stylistic choice—it’s a signature. An analysis of the genre distribution among the top three rising stars—Kimberly Snyder, Ping Tian, and Yang Zhao—reveals sharply contrasting approaches to musical identity and audience engagement.

```{r fig.width = 10, fig.height = 6}
#| code-fold: true
#| code-summary: "Show the code"
ggthemr("dust") 

top3_artists <- artist_summary %>%
  slice_max(rising_star_score, n = 3) %>%
  pull(artist_name)

genre_distribution <- influence %>%
  filter(artist_name %in% top3_artists) %>%
  count(artist_name, genre)

# Sort genre order
genre_order <- genre_distribution %>%
  count(genre, sort = TRUE) %>%
  pull(genre)

genre_distribution$genre <- factor(genre_distribution$genre, levels = genre_order)

# Create color palette (18 genres max)
genre_colors <- colorRampPalette(brewer.pal(8, "Dark2"))(length(genre_order))

# Calculate total per artist
genre_totals <- genre_distribution %>%
  group_by(artist_name) %>%
  summarise(total = sum(n))

# Plot
ggplot(genre_distribution, aes(x = reorder(artist_name, -n), y = n, fill = genre)) +
  geom_col(position = "stack", width = 0.7) +
  geom_text(data = genre_totals, aes(x = artist_name, y = total + 1, label = total),
            inherit.aes = FALSE, size = 5) +
  scale_fill_manual(values = genre_colors, name = "Genre") +
  labs(
    title = "Genre Distribution of Notable Works by the Top 3 Rising Stars",
    subtitle = "Comparing the creative range of Kimberly Snyder, Yang Zhao, and Ping Tian",
    caption = "Source: Oceanus Folk Archive  |  Chart by FU YILIN",
    x = NULL,
    y = "Number of Notable Works"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 8)),
    plot.title.position = "plot",
    plot.caption = element_text(size = 9, face = "italic", hjust = 1),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 9)
  ) +
  guides(fill = guide_legend(nrow = 3, byrow = TRUE))


```

**Kimberly Snyder** has carved out a niche rooted in Dream Pop, reinforcing a signature sound that leans into atmospheric melodies and introspective lyricism. Her body of notable works, while smaller in genre scope, speaks to focused artistic branding. With additional entries in Synthwave and Indie Rock, Snyder presents herself as a refined storyteller within a cohesive sonic world—one that appeals to a specific, loyal following drawn to dreamy soundscapes.

**Yang Zhao**, by contrast, showcases an emerging profile defined by adaptability. His notable works span a wider stylistic palette, from the earthy rhythms of Indie Folk to the brooding undercurrents of Doom Metal. This cross-genre presence suggests a rising artist who is as comfortable crafting intimate acoustic ballads as he is navigating experimental terrain. Zhao’s blend of raw emotionality and sonic risk-taking positions him as a boundary-pushing voice to watch.

Yet it is **Ping Tian** who stands out as the most prolific and genre-diverse of the trio. With the highest count of notable works among her peers, Tian’s creative output crosses the boundaries of Lo-Fi Electronica, Oceanus Folk, Avant-Garde Folk, and beyond. Her genre-spanning discography speaks to a bold and exploratory spirit, willing to disrupt musical norms while tapping into underground currents. This genre fluidity not only amplifies her reach but also marks her as a cultural chameleon—equally at home in experimental circles and broader listenerships.

### Annual Output Trend of Top 3 Rising Stars

The annual output chart reveals distinct artistic trajectories for Kimberly Snyder, Ping Tian, and Yang Zhao.

```{r fig.width=10, fig.height=6}
#| code-fold: true
#| code-summary: "Show the code"
ggthemr("dust")

# Step 1: Get top 3 artists
top3 <- artist_summary %>%
  slice_max(rising_star_score, n = 3) %>%
  pull(artist_name)

# Step 2: Prepare data with all years from first appearance
yearly_counts <- influence %>%
  filter(artist_name %in% top3) %>%
  count(artist_name, release_year) %>%
  complete(
    artist_name,
    release_year = seq(min(release_year, na.rm = TRUE), max(release_year, na.rm = TRUE)),
    fill = list(n = 0)
  )

# Step 3: Find peak years for labels
peaks <- yearly_counts %>%
  group_by(artist_name) %>%
  filter(n == max(n)) %>%
  slice(1) %>%
  ungroup()

# Step 4: Plot
ggplot(yearly_counts, aes(x = release_year, y = n, fill = artist_name)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_smooth(
    aes(color = artist_name, group = artist_name),
    method = "loess", se = FALSE, linewidth = 1, linetype = "dashed"
  ) +
  geom_text(
    data = peaks,
    aes(x = release_year, y = n + 0.5, label = paste0(n, " in ", release_year)),
    inherit.aes = FALSE,
    size = 3.8,
    fontface = "bold",
    color = "black"
  ) +
  scale_fill_manual(values = c(
    "Kimberly Snyder" = "#D55E00",
    "Ping Tian" = "darkgrey",
    "Yang Zhao" = "#0072B2"
  )) +
  scale_color_manual(values = c(
    "Kimberly Snyder" = "#D55E00",
    "Ping Tian" = "darkgrey",
    "Yang Zhao" = "#0072B2"
  )) +
  scale_x_continuous(breaks = seq(min(yearly_counts$release_year, na.rm = TRUE),
                                  max(yearly_counts$release_year, na.rm = TRUE), 2)) +
  labs(
    title = "Annual Output of the Top 3 Rising Stars",
    subtitle = "Tracking the yearly creative output of Kimberly Snyder, Ping Tian, and Yang Zhao",
    caption = "Source: Oceanus Folk Archive  |  Visualization by FU YILIN",
    x = "Year Released",
    y = "Number of Works",
    fill = "Artist",
    color = "Trend Line"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(face = "italic", size = 9, hjust = 1),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(
    fill = guide_legend(title.position = "top"),
    color = guide_legend(title.position = "top")
  )

```

Yang Zhao demonstrated an early surge in productivity, peaking with 6 works in 2013, followed by a gradual tapering in later years. In contrast, Ping Tian showed a steady and sustained rise beginning in the mid-2000s, reaching a modest peak of 2 works in 2009, and maintaining a consistent presence in subsequent years—suggesting long-term creative endurance. Kimberly Snyder emerged later but with strong momentum, hitting a peak of 5 works in 2023, indicating a more recent breakthrough and growing influence. The smoothed trend lines underscore these patterns: Yang Zhao’s curve shows early intensity and decline, Ping Tian’s trend is stable and gradual, while Kimberly Snyder’s trajectory is sharply upward—highlighting her as the most recent rising star.

### Annual Output Reveals the Rhythms of Three Rising Stars

```{r fig.width=10, fig.height=6}
#| code-fold: true
#| code-summary: "Show the code"
ggthemr("dust")

# Get top 3 artists
top3 <- artist_summary %>%
  slice_max(rising_star_score, n = 3) %>%
  pull(artist_name)

# Total annual output
total_output <- influence %>%
  filter(artist_name %in% top3) %>%
  count(artist_name, release_year, name = "total")

# Notable works per year
notable_output <- influence %>%
  filter(artist_name %in% top3, is_notable) %>%
  count(artist_name, release_year, name = "notable")

# Merge into one table
combined_output <- full_join(total_output, notable_output,
                             by = c("artist_name", "release_year")) %>%
  replace_na(list(total = 0, notable = 0))

# Plot: Bars = total, Dots = notable
ggplot(combined_output, aes(x = release_year, y = total)) +
  geom_col(fill = "#D55E00", width = 0.7) +
  geom_point(
    data = combined_output %>% filter(notable > 0),
    aes(x = release_year, y = notable),
    color = "black",
    size = 2.5
  ) +
  facet_wrap(~ artist_name, scales = "free_y") +
  labs(
    title = "Annual Output vs. Notable Works of Top 3 Rising Stars",
    subtitle = "Comparing release activity and charting success of Kimberly Snyder, Ping Tian, and Yang Zhao",
    caption = "Bars = Total works per year · Dots = Notable (Charting) works\nSource: Oceanus Folk Archive | Chart by FU YILIN",
    x = "Release Year",
    y = "Number of Works"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(size = 9, face = "italic", hjust = 1),
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank()
  )

```

**Yang Zhao** entered the spotlight with explosive force. His productivity peaked as early as 2013, when he released an impressive six works with 2 notable works, marking the height of his creative burst. But this intensity proved difficult to sustain. After his initial breakout, Zhao’s yearly output gradually tapered, hinting at a front-loaded career that may now be entering a quieter, more selective phase.

In contrast, **Ping Tian** presents the portrait of a slow-burn success. Her creative engine began building momentum in the mid-2000s, with a modest but steady cadence that culminated in a peak year of 2009. While she never experienced the dramatic surges seen in her peers’ careers, Tian’s enduring annual contributions signal a stable and reliable presence in the industry—an artist whose relevance is built on consistency rather than spikes.

**Kimberly Snyder**, the newest entrant among the three, tells a different story—one of rapid ascent. Her presence was scarcely noted in earlier years, but by 2023, she had achieved a remarkable five notable releases, revealing a sharp upward trajectory. Her trend line surges steeply, signaling not only a recent breakthrough but the potential for continued acceleration.

When viewed side by side, the trio’s trend lines tell a larger story: Zhao blazed brightly and early, Tian has weathered the years with quiet persistence, and Snyder is in the midst of a promising rise. Together, they sketch out the many paths to musical influence—from early fire to sustained flame to late bloom.

# **🎶 Chapter 3**

## Crafting the Soundscape

### A Comparative Look at Three Rising Influencers in Music Production

In an era where artistic versatility is as valued as vocal performance, a detailed look at the contribution network of three standout artists—Kimberly Snyder, Ping Tian, and Yang Zhao—reveals contrasting yet compelling creative trajectories.

```{r fig.width = 12, fig.height = 8}
#| code-fold: true
#| code-summary: "Show the code"

selected_artists <- c("Kimberly Snyder", "Ping Tian", "Yang Zhao")

#ID
artist_ids <- graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(name %in% selected_artists) %>%
  pull(id)


graph_selected_edges <- graph %>%
  activate(edges) %>%
  filter(
    `Edge Type` %in% c("ComposerOf", "ProducerOf", "LyricistOf") &
    (source %in% artist_ids | target %in% artist_ids)
  )


used_node_indices <- graph_selected_edges %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()


graph_selected <- graph_selected_edges %>%
  activate(nodes) %>%
  mutate(row_id = row_number()) %>%
  filter(row_id %in% used_node_indices) %>%
  select(-row_id)

# 计算影响力（节点大小）
graph_selected <- graph_selected %>%
  activate(nodes) %>%
  mutate(
    degree = centrality_degree(),
    notable_flag = ifelse(!is.na(notoriety_date), "Notable", "Other"),
    release_year = as.integer(release_date),
    label_with_year = case_when(
      `Node Type` %in% c("Song", "Album") & !is.na(release_year) ~ paste0(name, "\n(", release_year, ")"),
      TRUE ~ name
    ),
    size = case_when(
      `Node Type` == "Person" ~ degree * 1.5,
      !is.na(notoriety_date) ~ 6,
      TRUE ~ 2
    )
  )


ggraph(graph_selected, layout = "fr") +
  geom_edge_link(aes(color = `Edge Type`), alpha = 0.6, show.legend = TRUE) +
  geom_node_point(aes(color = `Node Type`, size = size, shape = notable_flag), show.legend = TRUE) +
  geom_node_text(aes(label = label_with_year), repel = TRUE, size = 3) +
  scale_edge_color_manual(values = c(
    ComposerOf = "#1f77b4",
    ProducerOf = "#2ca02c",
    LyricistOf = "#9467bd"
  )) +
  scale_color_manual(values = c(
    Person = "steelblue",
    Song = "gray40",
    Album = "tomato"
  )) +
  scale_shape_manual(values = c("Notable" = 17, "Other" = 16)) +
  guides(
    color = guide_legend(title = "Node Type"),
    shape = guide_legend(title = "Work Type"),
    edge_color = guide_legend(title = "Contribution")
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(size = 9, face = "italic", hjust = 1),
    legend.title = element_text(face = "bold", size = 11),
    legend.text = element_text(size = 10)
  ) +
  labs(
    title = "Artist Contribution Network with Influence and Release Year",
    subtitle = "Mapping the roles of Kimberly Snyder, Ping Tian, and Yang Zhao as composers, lyricists, and producers",
    caption = "Node size = contribution influence (degree centrality) · ▲ = notable work\nSource: Oceanus Folk Archive | Chart by FU YILIN"
  )
```

**Kimberly Snyder** emerges as the most prolific and critically acclaimed among the trio. Her dense web of connections across *ComposerOf*, *ProducerOf*, and *LyricistOf* roles speaks to both high output and strong creative control.

**Ping Tian** presents a different profile: measured, steady, and enduring. Her contributions span over three decades, beginning as early as `Bride’s Spectral Serenade (1997)` and extending to `Innocence Untouched (2030)`. While her network isn’t as dense, it shows remarkable consistency across all three creative roles—indicating a balanced career that has weathered shifts in musical trends while maintaining relevance.

**Yang Zhao**, by contrast, represents the energy of an emerging voice. Though his contribution network is smaller, it is rapidly expanding, with most of his work materializing post-2013. His role is strongest in lyricism and composition, with recent entries like `Symphonic Snowfall (2023)` and `Unadorned Silence (2027)` signaling a sharp rise in influence. His trajectory points not to breadth, but to a focused creative ascent, making him a name to watch as the next influential force in the music landscape.

Together, these three artists illustrate distinct paths to impact—Snyder through saturation and acclaim, Tian through longevity and balance, and Zhao through a focused surge of relevance in recent years.

### Mapping the Musical Influence of Three Visionary Artists

Zooming into broader influence to the music industry, the following visualized networks, built on referencing relationships such as *DirectlySamples*, *CoverOf*, *LyricalReferenceTo*, *InterpolatesFrom*, and *InStyleOf*, provide a nuanced lens into how creative DNA travels through time and genre.

```{r fig.width = 12, fig.height = 8}
#| code-fold: true
#| code-summary: "Show the code"
build_artist_plot <- function(artist_name_input) {
  # 1. Get artist ID
  artist_id <- graph %>%
    activate(nodes) %>%
    as_tibble() %>%
    filter(name == artist_name_input) %>%
    pull(id)

  # 2. Get works performed by the artist
  performed_work_ids <- graph %>%
    activate(edges) %>%
    filter(`Edge Type` == "PerformerOf", source == artist_id) %>%
    as_tibble() %>%
    pull(target)

  # 3. Get influence edges involving these works
  influence_types <- c("DirectlySamples", "LyricalReferenceTo", "CoverOf", "InterpolatesFrom", "InStyleOf")
  graph_influence_edges <- graph %>%
    activate(edges) %>%
    filter(`Edge Type` %in% influence_types,
           source %in% performed_work_ids | target %in% performed_work_ids)

  # 4. Get related node ids
  used_node_ids <- graph_influence_edges %>%
    activate(edges) %>%
    as_tibble() %>%
    select(from, to) %>%
    unlist() %>%
    unique()

  # 5. Subset the graph
  graph_artist_sub <- graph_influence_edges %>%
    activate(nodes) %>%
    mutate(row_id = row_number()) %>%
    filter(row_id %in% used_node_ids) %>%
    select(-row_id)

  # 6. Compute node attributes
  graph_artist_sub <- graph_artist_sub %>%
    activate(nodes) %>%
    mutate(
      degree = centrality_degree(),
      notable_flag = ifelse(!is.na(notoriety_date), "Notable", "Other"),
      release_year = as.integer(release_date),
      label_with_year = case_when(
        `Node Type` %in% c("Song", "Album") & !is.na(release_year) ~ paste0(name, "\n(", release_year, ")"),
        TRUE ~ name
      ),
      size = case_when(
        `Node Type` == "Person" ~ degree * 1.5,
        !is.na(notoriety_date) ~ 6,
        TRUE ~ 2
      )
    )

  # 7. Visualize
  ggraph(graph_artist_sub, layout = "fr") +
  geom_edge_link(aes(color = `Edge Type`), alpha = 0.6, show.legend = TRUE) +
  geom_node_point(aes(color = `Node Type`, size = size, shape = notable_flag), show.legend = TRUE) +
  geom_node_text(aes(label = label_with_year), repel = TRUE, size = 3) +
  scale_edge_color_manual(values = c(
    DirectlySamples = "#1f77b4",
    LyricalReferenceTo = "#9467bd",
    CoverOf = "#ff7f0e",
    InterpolatesFrom = "#2ca02c",
    InStyleOf = "#8c564b"
  )) +
  scale_color_manual(values = c(
    Person = "steelblue",
    Song = "gray40",
    Album = "tomato"
  )) +
  scale_shape_manual(values = c("Notable" = 17, "Other" = 16)) +
  guides(
    color = guide_legend(title = "Node Type"),
    shape = guide_legend(title = "Work Notability"),
    size = "none",
    edge_color = guide_legend(title = "Influence Type")
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(size = 9, face = "italic", hjust = 1),
    legend.title = element_text(face = "bold", size = 11),
    legend.text = element_text(size = 10)
  ) +
  labs(
    title = paste("Influence Network of", artist_name_input),
    subtitle = "Mapped through song references: covers, lyrical callbacks, sampling, stylistic echoes, and interpolations",
    caption = "▲ = notable work · Node size = influence degree · Source: Oceanus Folk Archive | Chart by FU YILIN"
  )
}
```

#### **Kimberly Snyder**

```{r fig.width = 12, fig.height = 8}
#| code-fold: true
#| code-summary: "Show the code"
build_artist_plot("Kimberly Snyder")

```

Kimberly Snyder’s influence network is a dynamic hub of interconnectivity. Works like `Beethoven’s Unheard Conversation (2024)` and `Echoes of a Global Generation (2023)` serve as influential anchors, receiving numerous stylistic nods from other artists. Her music is not only frequently covered or sampled but often becomes the stylistic blueprint for future compositions. The presence of multiple notable hits (▲) and the centrality of her network nodes signal that Snyder is both a trendsetter and a continuous source of musical evolution. Her influence arcs outward and forward, with references extending as far as `Farewell to the Fading Sun (2035)`, underscoring a legacy in the making.

#### **Ping Tian**

```{r fig.width = 10, fig.height = 6}
#| code-fold: true
#| code-summary: "Show the code"
build_artist_plot("Ping Tian")

```

Ping Tian, on the other hand, paints a portrait of quiet resonance and long-term relevance. Her influence network, though less clustered than Snyder’s, stretches across a remarkable timeline—from `Bride’s Spectral Serenade (1997)` to 2030’s `Innocence Untouched`. She inspires through lyrical motifs and emotional tone, with works like `Lift Me Higher (2012)` and `Back Up Off Me (2019)` frequently echoed via interpolations and references. Rather than dominating stylistic discourse, Tian infuses it—her music acting as connective tissue across decades of songwriting. Her contribution is subtle but foundational.

#### **Yang Zhao**

```{r fig.width = 10, fig.height = 6}
#| code-fold: true
#| code-summary: "Show the code"
build_artist_plot("Yang Zhao")

```

Meanwhile, Yang Zhao’s influence is precise and potent. At the heart of his network is `Little Alfie’s Lament (2013)`—a standout hit that catalyzes a wave of inspired works. His referencing graph, though less sprawling, is rich in directional influence. Songs like `Eva’s Revolutionary Stride (2030)` and `After the Last Tear Falls (2020)` point to Zhao’s growing impact in recent years. The graph’s temporal reach, spanning from early 2000s references to futuristic echoes in the 2030s, positions him as an emerging artist whose emotional narrative and stylistic clarity are shaping the next generation.

# **🎶 Chapter 4**

## **Rising Waves in Oceanus Folk**

### **Who Will Shape the Genre Next?**

As Oceanus Folk continues to captivate audiences with its emotive melodies and coastal rhythms, a new generation of talent is beginning to emerge. Our team has applied a multi-layered algorithm to predict the next big names likely to rise to stardom in the genre over the next five years.

#### Data-Driven Discovery

The predictive model focuses exclusively on artists who have released Oceanus Folk works since 2030, eliminating outliers like long-established superstar Sailor Shift to surface new potential. The analysis considers several key metrics:

-   Total output: How many works the artist has produced since 2030.

-   Chart impact: A double-weighted score for works that have appeared on notoriety charts.

-   Career momentum: The year of their first release and first notable appearance help determine their early trajectory.

-   Multi-role versatility: Artists who serve not just as performers but also as producers, composers, or lyricists receive bonus credit for creative range.

Each artist is assigned a “potential score” combining these elements. This blended measure allows the model to highlight artists not only making waves today but building the foundation for sustained impact.

### The Next 3 to Watch

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Step 1: Identify artists involved in multiple roles
multi_role_artists <- edges_tbl %>%
  filter(`Edge Type` %in% c("PerformerOf", "ProducerOf", "ComposerOf", "LyricistOf")) %>%
  count(source) %>%
  filter(n >= 2) %>%
  pull(source)

# Step 2: Summarize next-gen rising stars
next_stars <- influence %>%
  filter(
    genre == "Oceanus Folk",
    release_year >= 2030,
    artist_name != "Sailor Shift"
  ) %>%
  group_by(artist_name) %>%
  summarise(
    total_works = n(),
    notable_hits = sum(is_notable),
    first_release = min(release_year, na.rm = TRUE),
    last_release = max(release_year, na.rm = TRUE),
    first_notoriety = if (all(is.na(notoriety_year))) NA_integer_ else min(notoriety_year, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    years = paste(first_release, "–", last_release),
    is_multi_role = artist_name %in% (nodes_tbl %>%
                                        filter(id %in% multi_role_artists) %>%
                                        pull(name)),
    active_years = last_release - first_release + 1,
    potential_score = total_works + notable_hits * 2 + ifelse(is_multi_role, 2, 0)
  ) %>%
  arrange(desc(potential_score)) %>%
  slice_head(n = 3) %>%
  mutate(
    profile_md = glue::glue(
      "::: {{.panel .panel-borderless}}\n",
      "**{artist_name}**\n\n",
      "- **Active Years:** {years}\n",
      "- **Total Works:** {total_works}\n",
      "- **Notable Works:** {notable_hits}\n",
      "- **Multi-role Contributor:** {ifelse(is_multi_role, 'Yes', 'No')}\n",
      "- **Career Span:** {active_years} year(s)\n",
      ":::"
    )
  )


```

```{r echo=FALSE, results='asis'}
cat(paste(next_stars$profile_md, collapse = "\n\n"))

```

Their profiles suggest not just short-term success, but long-term influence. These rising voices are poised to define what Oceanus Folk will sound like in the years to come.

```{r fig.width = 10, fig.height = 6}
#| code-fold: true
#| code-summary: "Show the code"
# Step 1: Get top 3 artist names
top3_names <- next_stars$artist_name

# Step 2: Get performer edges for top 3
artist_song_edges <- edges_tbl %>%
  filter(`Edge Type` == "PerformerOf") %>%
  semi_join(nodes_tbl %>% filter(name %in% top3_names), by = c("source" = "id")) %>%
  select(artist_id = source, song_id = target)

# Step 3: Extract nodes
artist_nodes <- nodes_tbl %>%
  filter(name %in% top3_names) %>%
  mutate(label = name, type = "Artist")

song_nodes <- nodes_tbl %>%
  filter(id %in% artist_song_edges$song_id) %>%
  mutate(label = name, type = "Song")

# Step 4: Combine and re-index nodes
all_nodes <- bind_rows(artist_nodes, song_nodes) %>%
  mutate(graph_id = row_number())

# Step 5: Map original IDs to graph IDs
node_map <- all_nodes %>% select(id, graph_id)

edges_mapped <- artist_song_edges %>%
  left_join(node_map, by = c("artist_id" = "id")) %>%
  rename(from_id = graph_id) %>%
  left_join(node_map, by = c("song_id" = "id")) %>%
  rename(to_id = graph_id) %>%
  select(from = from_id, to = to_id)

# Step 6: Create graph and plot
song_graph <- tbl_graph(nodes = all_nodes, edges = edges_mapped, directed = FALSE)

ggraph(song_graph, layout = "fr") +
  geom_edge_link(color = "gray80", alpha = 0.6) +
  geom_node_point(aes(color = type, shape = type, size = type)) +
  geom_node_text(aes(label = label), repel = TRUE, size = 3.2) +
  scale_color_manual(values = c("Artist" = "#D55E00", "Song" = "steelblue")) +
  scale_shape_manual(values = c("Artist" = 17, "Song" = 16)) +
  scale_size_manual(values = c("Artist" = 6, "Song" = 3)) +
  labs(
    title = "Top 3 Rising Stars and Their Performed Works",
    subtitle = "Network visualization of artist-song relationships",
    caption = "Source: Oceanus Folk Archive | Chart by FU YILIN"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5),
    legend.position = "bottom"
  )

```

In a striking detail from the network graph, two songs — Tide-Worn Tales and The Sunken Compass — appear anchored at the center of the visual, bridging multiple artist clusters. This suggests that both pieces were performed by more than one rising star, serving as collaborative or cross-influenced works in the Oceanus Folk scene.
